---
title: "index"
author: "Julien BARRERE"
date: "5 octobre 2021"
output: 
  html_document:
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "plots/fig_")
```

# Packages and set up

We start by loading the required packages. `targets` must be version 0.5.0.9000 or above. 

```{r, eval = FALSE}
require(c("dplyr", "ggplot2", "targets", "tidyr", "RColorBrewer", "lme4",
          "data.table", "knitr", "stringr", "measurements", "sf", "raster",
          "rgdal", "exactextractr", "rgeos", "rnaturalearth", "rnaturalearthdata", 
          "ggspatial", "cowplot", "rjags", "coda", "R2jags", "MASS", "ggmcmc"))
```

Second, we call the `targets` library. The `tar_unscript()` R function removes the `_targets_r` directory previously written by non-interactive runs of the report. It prevents the pipeline from containing superfluous targets.

```{r}
library(targets)
tar_unscript()
```

# Globals

We define some global options/functions common to all targets, and load the functions located in the R directory of the repository. 

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
options(stringsAsFactors=FALSE)
source("R/functions_data.R")
source("R/functions_plot.R")
source("R/functions_analysis.R")
tar_option_set(packages = c("dplyr", "ggplot2", "targets", "tidyr", "RColorBrewer", "lme4",
          "data.table", "knitr", "stringr", "measurements", "sf", "raster",
          "rgdal", "exactextractr", "rgeos", "rnaturalearth", "rnaturalearthdata", 
          "ggspatial", "cowplot", "rjags", "coda", "R2jags", "MASS", "ggmcmc"))
```


# Importation of french NFI data

We start by importing the following data on: 

- NFI plots (`NFI_plot`) and associated ecological data (`NFI_ecological_data`) and elevation (`NFI_plot_elevation`)
- NFI individual trees that were either alive (`NFI_tree_alive`) or dead (`NFI_tree_dead`) 
- NFI plots remeasured (`NFI_plot_remeasure`) 
- Individual trees remeasured that were either alive  (`NFI_tree_alive_remeasure`) or dead (`NFI_tree_dead_remeasure`) at first measurement. 

```{targets import-NFI-data}
list(
  tar_target(NFI_plot, read_plot(path = "data/FrenchNFI")), 
  tar_target(NFI_plot_elevation, read.csv("data/FrenchNFI/elevation_NFIplots.csv", sep = ";")),
  tar_target(NFI_ecological_data, read_ecological_data(path = "data/FrenchNFI")),
  tar_target(NFI_tree_alive, read_tree(path = "data/FrenchNFI")), 
  tar_target(NFI_tree_dead, read_dead_tree(path = "data/FrenchNFI")), 
  tar_target(NFI_plot_remeasure, read_reameasure_plot(path = "data/FrenchNFI_remeasure")), 
  tar_target(NFI_tree_alive_remeasure, read.csv("data/donnees_arbres_retour/foret_c135.csv", sep = ";")), 
  tar_target(NFI_tree_dead_remeasure, read.csv("data/donnees_arbres_retour/foret_morts.csv", sep = ";"))
)

```

Second, we merge : 

* the data on alive and dead trees to obtain `NFI_tree`
* the remeasures of previously alive and dead trees to obtain `NFI_tree_remeasured`.

```{targets NFI_tree_join}
list(
  tar_target(NFI_tree, merge_dead_alive(NFI_tree_alive, NFI_tree_dead)),
  tar_target(NFI_tree_remeasured, 
             merge_NFI_remeasured(NFI_tree_alive_remeasure, NFI_tree_dead_remeasure))
)
```


# Formating of French NFI data for the TreeMort project

## Species data

We start by importing :

* The existing species table in the NFI directory (`NFI_species`).
* A correspondence table linking genus and families. This table was built from the NCBI Taxonomy database ([Federhen 2012](https://doi.org/10.1093/nar/gkr1178)).

```{targets import-species-NFI}
list(
  tar_target(NFI_species, read.csv("data/FrenchNFI/species.csv", sep = ",")), 
  tar_target(NFI_genus_species_correspondence, read.csv("data/TreeMort/correspondanceGenusFamilyNFI.csv", sep = ";"))
)
```

Using these two tables, we create a species table that matches TreeMort template (`TreeMort_species`), and write this table in the TreeMort directory. 

```{targets format-species-NFI-to-TreeMort}
tar_target(TreeMort_species, Format_species_TreeMort(NFI_species, NFI_genus_species_correspondence))
```




## Trees data

We start by formatting the table of the first (census 1) and second (census 2) measurements (both dead and alive trees).

```{targets format-trees-NFI-to-TreeMort-by-census}
list(
  tar_target(TreeMort_tree_census1, Format_trees_census1_TreeMort(NFI_tree, TreeMort_species)),
  tar_target(TreeMort_tree_census2, Format_trees_census2_TreeMort(NFI_tree_remeasured, TreeMort_tree_census1))
)
```

We merge these two dataframes to obtain the complete tree NFI dataset formatted for TreeMort (`TreeMort_tree`), that we write in the TreeMort directory. 

```{targets format-trees-NFI-to-TreeMort}
tar_target(TreeMort_tree, 
           rbind(subset(TreeMort_tree_census1, 
                        tree.id %in% TreeMort_tree_census2$tree.id),
                 TreeMort_tree_census2))
```

## Plots data

We format the `NFI_plot` table to fit into TreeMort template, and we write the formated table (`TreeMort_plot`) in the TreeMort directory.

```{targets format-plots-NFI-to-TreeMort}
list(
  tar_target(NFI_stand_age, Compute_NFI_stand_age(NFI_tree)), 
  tar_target(TreeMort_plot, Format_plots_TreeMort(NFI_plot, NFI_ecological_data, NFI_plot_elevation, NFI_stand_age))
)
```


## Census data

First, we need to create a management variable describing the occurence of a cut in the last 5 years for each census. 

```{targets create-management-census}
tar_target(NFI_census_management, compute_management_census(NFI_plot, NFI_plot_remeasure))
```

We format the `TreeMort_tree` table to create a census table fitting TreeMort template, and we write the formated table (`TreeMort_census`) in the TreeMort directory.

```{targets format-census-TreeMort}
tar_target(TreeMort_census, Format_census_TreeMort(TreeMort_tree, NFI_census_management))
```

## Meta data

We write a metadata table in the TreeMort directory to specify important information on some columns. 

```{targets metadata-TreeMort}
tar_target(TreeMort_metadata, Meta_data_TreeMort())
```

# Importation and formating of FUNDIV data

## Importation

We start by importing the different datatables of FUNDIV in their original format. 
```{targets import-FUNDIV}
list(
  tar_target(FUNDIV_tree_original_NoFR, read.csv(file.path("data/FUNDIV", "FunDiv_trees_Nadja.csv"),
                                                 stringsAsFactors=FALSE) %>% dplyr::filter(country != "FG")),
  tar_target(FUNDIV_species, read.csv(file.path("data/FUNDIV", "FunDiv_species_Nadja.csv"),
                                     stringsAsFactors=FALSE, fileEncoding = "cp1252")), 
  tar_target(FUNDIV_plots_original_noFR, read.csv(file.path("data/FUNDIV", "FunDiv_plots_Nadja.csv"),
                                                  stringsAsFactors=FALSE) %>% dplyr::filter(country != "FG")),
  tar_target(FUNDIV_climate, read.csv(file.path("data/FUNDIV", "FunDiv_plots_climate_sapropos.csv"),
                                      stringsAsFactors=FALSE)),
  tar_target(FUNDIV_management, read.csv(file.path("data/FUNDIV", "FunDivEUROPE_plot_management.csv"),
                                      stringsAsFactors=FALSE))
)
```

## Inclusion of French NFI remeasured data

Using French NFI data formatted for TreeMort, we format the French tree data with true mortality observations to FUNDIV template. 

```{targets format-trees-TreeMort-to-FUNDIV}
tar_target(FUNDIV_tree_original_FR, Format_trees_TreeMort_to_FUNDIV(TreeMort_tree, FUNDIV_species))
```

We do the same with the french plot data: 

```{targets format-plots-TreeMort-to-FUNDIV}
tar_target(FUNDIV_plots_original_FR, 
           Format_plots_TreeMort_to_FUNDIV(NFI_ecological_data, TreeMort_plot, FUNDIV_tree_original_FR))
```

Lastly, we can replace the datasets with and without french data for trees and plots. We also apply a small correction to the weight in the tree table

```{targets merge-FrenchNFI-FUNDIV}
list(
  tar_target(FUNDIV_tree_original, 
             correct_weight_FUNDIV(rbind(FUNDIV_tree_original_NoFR, FUNDIV_tree_original_FR))), 
  tar_target(FUNDIV_plots_original, 
             rbind(FUNDIV_plots_original_noFR, FUNDIV_plots_original_FR))
)
```

We format the original FUNDIV data to better fit the IPM (and to include climatic variables). 

```{targets format-IPM-FUNDIV}
tar_target(FUNDIV_tree, 
           read_FUNDIV_tree_data(FUNDIV_tree_original, FUNDIV_plots_original, FUNDIV_species, 
                              FUNDIV_climate, FUNDIV_management, remove_harv = F))
```

## Formating to Finnish NFI template

The Finnish NFI template is slightly different from the French NFI template or FUNDIV template. Therefore, we need to format them. 

```{targets format-FUNDIV-to-FinnishNFI}
tar_target(FinnishNFI_French_tree, 
           Format_trees_FUNDIV_to_FinnishNFI(subset(FUNDIV_tree, country == "FR"), NFI_tree, 
                                              NFI_plot_elevation, NFI_plot_remeasure, 
                                              NFI_ecological_data))
```

We write a metadata table in the FinnishNFI directory to specify important information on some columns. 

```{targets metadata-FinnishNFI}
tar_target(FinnishNFI_metadata, Meta_data_FinnishNFI())
```


## Inclusion of disturbance data in FUNDIV

We create a circular buffer around each FUNDIV plot center, and we associate each plot to the percentage of the buffer area covered by each type of disturbance ([Senf & Seidl, 2021a](https://doi.org/10.1111/gcb.15679)) that occured between 1986 and 2020. We use three size of circular buffer: 15m, 100m and 200m: 

```{targets import-disturbance-Senf-areaPercentage}
list(tar_target(disturbance_per_plot_15m, 
           Get_disturbance_per_plot(15, FUNDIV_tree, "data/Disturbance")), 
     tar_target(disturbance_per_plot_100m, 
           Get_disturbance_per_plot(100, FUNDIV_tree, "data/Disturbance")), 
     tar_target(disturbance_per_plot_200m, 
           Get_disturbance_per_plot(200, FUNDIV_tree, "data/Disturbance")))
```

Secondly, we add this information in the FUNDIV database.

```{targets add-Senf-areaPercentage-to-FUNDIV}
list(tar_target(FUNDIV_tree_disturbance15m, 
           add_disturbance_to_FUNDIV(FUNDIV_tree, disturbance_per_plot_15m)), 
     tar_target(FUNDIV_tree_disturbance100m, 
           add_disturbance_to_FUNDIV(FUNDIV_tree, disturbance_per_plot_100m)), 
     tar_target(FUNDIV_tree_disturbance200m, 
           add_disturbance_to_FUNDIV(FUNDIV_tree, disturbance_per_plot_200m)))
```

The code below extracts the prevalence (% of buffer area of all plots covered) of each disturbance type per country and per year, in order to monitor temporal variations in disturbance regimes. 

```{targets get-disturbance-prevalence-per-country-and-year}
tar_target(prevalence_per_country_per_year, 
           get_annualPrevalence(disturbance_per_plot_200m, FUNDIV_tree))
```




# Fit of bayesian mortality model

The code below format French NFI data to fit the bayesian mortality model presented in [this document](https://www.overleaf.com/project/61d5583436dde4fa2f5cec66) in JAGS for several tree species, with the french data only. Two options: either by fitting the occurrence of a disturbance (Dj) using Senf data, either by using true data on the occurrence of a disturbance. 

```{targets format-data-mortality-France}
list(
  tar_target(data_jags_Aalba, 
           format_FrenchNFI_mortality(FUNDIV_tree, FUNDIV_climate, 
                                      disturbance_per_plot_200m, 
                                      NFI_plot_remeasure, "Abies alba")), 
  tar_target(data_jags_Aalba_2, 
           format_FrenchNFI_mortality_2(FUNDIV_tree, FUNDIV_climate, 
                                      disturbance_per_plot_200m, 
                                      NFI_plot_remeasure, "Abies alba"))
)
```

From these formatted data sets, the code below generate simulated data also adapted to the Bayesian model
```{targets generate-data-mortality-France}
list(
  tar_target(data_jags_generated, 
           generate_data_jags(10, 1000, data_jags_Aalba)), 
  tar_target(data_jags_generated_2, 
           generate_data_jags_2(10, 1000, data_jags_Aalba_2))
)
```

Lastly, we can fit the mortality model with both true and simulated data. 
```{targets fit-mortality-France}
list(
  tar_target(jags_simulated, 
             fit_mortality(data_jags_generated$data, n.chains = 3, n.iter = 5000, n.burn = 1000, n.thin = 1)), 
  tar_target(jags_simulated_2, 
             fit_mortality_2(data_jags_generated_2$data, n.chains = 3, n.iter = 5000, n.burn = 3000, n.thin = 1)))

```


# Make all plots, run pipeline and write tables

```{targets make-all-plots}
list(
  tar_target(Map_areaCoveredPerDisturbance_FUNDIVplots, 
           Map_FUNDIVplots_areaCoveredPerDisturbance(FUNDIV_tree_disturbance100m)), 
  tar_target(plot_Compare_NFI_Senf_disturbance_area, 
           Compare_NFI_Senf_disturbance_area(FUNDIV_tree_disturbance100m, NFI_plot_remeasure)), 
  tar_target(plot_prevalence_per_country_per_year, 
           plot_annualPrevalence(prevalence_per_country_per_year)), 
  tar_target(plot_disturbedArea_perMortalityRate_15m, 
           plot_areaDisturbance_perMortalityRate(FUNDIV_tree_disturbance15m, 
                                                 "Buffer = 15m radius", 
                                                 death.in = "natural mortality")), 
  tar_target(plot_disturbedArea_perMortalityRate_100m, 
           plot_areaDisturbance_perMortalityRate(FUNDIV_tree_disturbance100m, 
                                                 "Buffer = 100m radius", 
                                                 death.in = "natural mortality")), 
  tar_target(plot_disturbedArea_perMortalityRate_200m, 
           plot_areaDisturbance_perMortalityRate(FUNDIV_tree_disturbance200m, 
                                                 "Buffer = 200m radius", 
                                                 death.in = "natural mortality")), 
  tar_target(plot_disturbedArea_perMortalityRate_200m_harvest, 
           plot_areaDisturbance_perMortalityRate(FUNDIV_tree_disturbance200m, "", 
                                                 death.in = "harvested")), 
  tar_target(plot_disturbedArea_perMortalityRate_200m_unknown, 
           plot_areaDisturbance_perMortalityRate(FUNDIV_tree_disturbance200m, "", 
                                                 death.in = "unknown cause")), 
  tar_target(plot_disturbedArea_perMortalityRate_200m_alldeath, 
           plot_areaDisturbance_perMortalityRate(FUNDIV_tree_disturbance200m, "", 
                                                 death.in = c("natural mortality", "harvested", "unknown cause"))), 
  tar_target(plot_compare_mortality_tree_plot_level, 
             plot_mortality_tree_plot(NFI_tree_alive_remeasure, NFI_plot_remeasure)),
  tar_target(plot_probability_to_be_harvested, 
             plot_harvest_probability(NFI_tree_alive_remeasure, NFI_plot_remeasure, NFI_tree_alive)),
  tar_target(plot_probability_to_be_harvested2, 
             plot_harvest_probability2(NFI_tree_alive_remeasure, NFI_plot_remeasure, NFI_tree_alive)), 
  tar_target(fig_status_disturbance, 
             plot_status_disturbance(NFI_tree, NFI_tree_remeasured, NFI_plot_remeasure, NFI_species)),
  tar_target(fig_convergence_simulated, plot_convergence(jags_simulated, "Simulated data - Dj latent")), 
  tar_target(fig_convergence_simulated_2, plot_convergence(jags_simulated_2, "Simulated data - Dj true data")), 
  tar_target(fig_compare_jags_simulated, plot_compare_jags_simulated(data_jags_generated, jags_simulated)), 
  tar_target(fig_compare_jags_simulated_2, plot_compare_jags_simulated(data_jags_generated_2, jags_simulated_2))
)
```

```{r, message=TRUE, results='hide'}
tar_make()
```

```{r}
# Create output directories
if(!dir.exists("output")) dir.create("output")
if(!dir.exists("output/TreeMort")) dir.create("output/TreeMort")
if(!dir.exists("output/FUNDIV")) dir.create("output/FUNDIV")
if(!dir.exists("output/FinnishNFI")) dir.create("output/FinnishNFI")
write.table(tar_read(TreeMort_species), 
          file = "output/TreeMort/spp_data.csv", row.names = F)
write.table(tar_read(TreeMort_tree), 
          file = "output/TreeMort/tree_data.csv", row.names = F)
write.table(tar_read(TreeMort_plot), 
          file = "output/TreeMort/plot_data.csv", row.names = F)
write.table(tar_read(TreeMort_census), 
          file = "output/TreeMort/census_data.csv", row.names = F)
write.table(tar_read(TreeMort_metadata), 
          file = "output/TreeMort/meta_data.csv", row.names = F)
write.table(tar_read(FUNDIV_tree_original), 
          file = "output/FUNDIV/FUNDIV_tree_original.csv", row.names = F)
write.table(tar_read(FUNDIV_plots_original), 
          file = "output/FUNDIV/FUNDIV_plots_original.csv", row.names = F)
write.table(tar_read(FinnishNFI_French_tree), 
          file = "output/FinnishNFI/FinnishNFI_French_tree.csv", row.names = F)
write.table(tar_read(FinnishNFI_metadata), 
          file = "output/FinnishNFI/FinnishNFI_meta_data.csv", row.names = F)
```

# Data exploration - Disturbance data and FUNDIV

We plot the percentage of the buffer around each plot covered by the three types of disturbance. 
```{r AreaCoveredPerDisturbance, fig.height=3, fig.width=10, fig.align='center', dpi=600}
tar_read(Map_areaCoveredPerDisturbance_FUNDIVplots)
```



In the plot below, we compute from the dataset `disturbance_per_plot_200m` the percentage of the buffer area affected by each disturbance for each year and each country, and show this temporal and spatial variability. We also added the year of first and last measurement for each country (vertical dotted black lines). 

```{r plot-prevalence-per-country-per-year, fig.height=5, fig.width=10, fig.align='center', dpi=600}
tar_read(plot_prevalence_per_country_per_year)
```


To make the plots below, we computed the mortality rate of each plot (harvested trees and recruits were removed), converted it into a qualitative variable and we computed for each class the percentage of area covered by each type of disturbance. We make one plot per buffer size.  

```{r plot-disturbedArea-perMortalityRate-15m, fig.height=8, fig.width=10, fig.align='center', dpi=600}
cowplot::plot_grid(tar_read(plot_disturbedArea_perMortalityRate_15m) + 
                     ggplot2::theme(legend.position = "none"), 
          tar_read(plot_disturbedArea_perMortalityRate_100m) + 
            ggplot2::theme(legend.position = "none"), 
          tar_read(plot_disturbedArea_perMortalityRate_200m) + 
            ggplot2::theme(legend.position = "none"), 
          cowplot::get_legend(tar_read(plot_disturbedArea_perMortalityRate_200m)), 
          align = c("h", "v"), nrow = 2)

```

In the 3 plots above, we only considered natural mortality when calculating mortality rates (harvested trees, or trees that died of unknown causes were removed from the dataset). In the three plots below, we show the same figure than above (with 200m-radius buffer), but with different approaches to calculate the mortality rates (with harvest only, unknown cause only, or all cause of death confounded)

```{r plot-disturbedArea-perMortalityRate-200m-harvest, fig.height=8, fig.width=10, fig.align='center', dpi=600}
cowplot::plot_grid(
  tar_read(plot_disturbedArea_perMortalityRate_200m) + ggplot2::theme(legend.position = "none") + 
    ggplot2::ggtitle(""), 
  tar_read(plot_disturbedArea_perMortalityRate_200m_harvest) + ggplot2::theme(legend.position = "none"), 
  cowplot::get_legend(tar_read(plot_disturbedArea_perMortalityRate_200m_alldeath)), 
  tar_read(plot_disturbedArea_perMortalityRate_200m_unknown) + ggplot2::theme(legend.position = "none"), 
  tar_read(plot_disturbedArea_perMortalityRate_200m_alldeath) + 
    ggplot2::theme(legend.position = "none", axis.title.x = ggplot2::element_text(size = 10)), 
  rel_widths = c(1, 1, 0.3), nrow = 2)
```


The plot below shows how the tree status (percentage of the basal area) changes with the occurence, nature and severity of a disturbance at the plot level (based on French NFI tru estimation of disturbance). 

```{r fig-status-disturbance-FrenchNFI, fig.height=5, fig.width=7, fig.align='center', dpi=600}
tar_read(fig_status_disturbance)
```





# Data analysis - Disturbance data and FUNDIV

To take into account salvage logging in the survival model, we need to compute a probability to be harvested knowing that the tree is alive, or dead. To that end, we can use the estimation by NFI agents in France of the disturbance that occurred in the plot. First, we need to compare the mortality at plot level with real mortality at tree level, computed with two different approaches (with or without mortality rate).  


```{r plot-compare-mortality-tree-plot-level, fig.height=5, fig.width=10, fig.align='center', dpi=600}
tar_read(plot_compare_mortality_tree_plot_level)
```

Mortality rate at the tree level matches mortality at the plot level when it is computed with the harvest rate. From this result, we can assume that in a plot affected by a disturbance, all trees that were harvested were dead before to be harvested. In the case of undisturbed, it is not possible to know whether harvested trees experienced background mortality before, or were alive. Therefore, we need to make the assumption that trees that died from background mortality have the same probability to be harvested than alive trees. 

Based on these assumptions, we modeled the harvest probability with a simple binomial glm with a logit link function as a function of the status (alive or dead due to disturbance or background mortality) and of the dbh at census 1. 


```{r plot-harvest-probability, fig.height=6.4, fig.width=8, fig.align='center', dpi=600}
tar_read(plot_probability_to_be_harvested)
```


```{r plot-harvest-probability-two, fig.height=4, fig.width=10, fig.align='center', dpi=600}
tar_read(plot_probability_to_be_harvested2)
```


# Data analysis - Bayesian mortality model

## With Dj estimated as a latent variable

Convergence:
```{r plot-convergence-simulated, fig.height=6, fig.width=8, fig.align='center', dpi=600}
tar_read(fig_convergence_simulated)
```

Comparison of true simulated parameter value and fitted values. 
```{r plot-compare-parameters-simulated, fig.height=6, fig.width=8, fig.align='center', dpi=600}
tar_read(fig_compare_jags_simulated)
```

## With true data on Dj

Convergence:
```{r plot-convergence-simulated-2, fig.height=6, fig.width=8, fig.align='center', dpi=600}
tar_read(fig_convergence_simulated_2)
```

Comparison of true simulated parameter value and fitted values. 
```{r plot-compare-parameters-simulated-2, fig.height=6, fig.width=8, fig.align='center', dpi=600}
tar_read(fig_compare_jags_simulated_2)
```

